---
title: "PRISM: PROGRAMMABLE INTERFACE FOR STATISTICAL & SIMULATION MODELS"
site: "bookdown::bookdown_site"
runtime: "shiny"
output:
  bookdown::gitbook:
    lib_dir: "book_assets"
    delete_merged_file: TRUE
    table_css: FALSE
    output_dir: "docs"
    css: ["pretty-styles.css", "cayman.css", "python.css", "index.html"]
    js: "js/list-group.js"
    config:
      toc:
        collapse: subsection
        scroll_highlight: yes
        before: null
        after: null
      toolbar:
        position: fixed
      edit : null
      download: null
      search: yes
      fontsettings:
        theme: white
        family: sans
        size: 2
      sharing:
        facebook: yes
        twitter: yes
        google: no
        linkedin: no
        weibo: no
        instapaper: no
        vk: no
        all: ['facebook', 'google', 'twitter', 'linkedin', 'weibo', 'instapaper']
  bookdown::pdf_book:
    keep_tex: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ">"
)
```

```{r, echo = FALSE, eval = TRUE}
themeColors = c("#330033", "#8cf2f2", "#c51672", "#007bff")
```
# Introduction {#introduction}

PRISM is a platform that makes decision-analytics and clinical prediction models accessible on the cloud. PRISM enables end-users to interact with the model they are interested in, without worrying about having the required hardware resources, or setting up complex code on their computer.

In its current implementation, PRISM can host models created in R, and can be accessed through Microsoft Excel and R.

[PRISM Introduction](http://resp.core.ubc.ca/research/Specific_Projects/PRISM).


