
# Application Programming Interface (API) {#api}

This section is geared towards people who are less familiar with APIs, or who want a refresher. If you are comfortable with APIs and don't need a review, you can skip to the next section.

## What is an API? {-}

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("user-server-api.html"), collapse="\n")
HTML(rawHTML)
```

## REST {-}

REST stands for **RE**presentational **S**tate **T**ransfer. Ah yes, clear as mud. I understand the words individually, but...

REST is basically the rules for how the client (your computer) can talk to the server (the computer with PRISM, say). It is the rules for the API.

REST is a simple language. It isn't like English, or French, or Mandarin. It only has 4 words! So much easier to be bilingual...

The 4 words are: **GET**, **POST**, **PUT**, and **DELETE**. Simple, right? It turns out, the difficulty comes in defining the words. The official definitions are fairly broad and rather vague, when in practice the use of the words depend on: a) context and b) general rules of thumb. A bit like English after all!


## Diving in {-}

I have found it is often easiest to learn by example, so I'm going to walk through an example of how to use our PRISM API. If you don't have much programming experience, or if you want to use Excel, I would recommend doing the [API in Excel](#api-in-excel). Otherwise, you can choose your favourite programming language to get started:

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("link-buttons.html"), collapse="\n")
HTML(rawHTML)
```

