
# Application Programming Interface (API) {#api}

This section is geared towards people who are less familiar with APIs, or who want a refresher. If you are comfortable with APIs and don't need a review, you can skip to the next section.

## What is an API? {-}

```{r api-graphic, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("user-server-api.html"), collapse="\n")
HTML(rawHTML)
```

## REST {-}

REST stands for **RE**presentational **S**tate **T**ransfer. Ah yes, clear as mud. I understand the words individually, but...

REST is basically the rules for how the client (your computer) can talk to the server (the computer with PRISM, say). It is the rules for the API.

REST is a simple language. It isn't like English, or French, or Mandarin. It only has 4 words! So much easier to be bilingual...

The 4 words are: **GET**, **POST**, **PUT**, and **DELETE**. Simple, right? It turns out, the difficulty comes in defining the words. The official definitions are fairly broad and rather vague, when in practice the use of the words depend on: a) context and b) general rules of thumb. A bit like English after all!

### GET {-}

When using an API, you have 2 options: you may send some data, and receive something back, or you can just receive something back. **GET** is the function where you just receive data. For this tutorial, we are going to use the Google Books API, which I chose because it has good documentation and doesn't require you to get an API key.

Google Books is a huge database of books. The database lives on a computer somewhere else, called the Server. You, at home on your computer (Client), would like to see if you can find a book using Google's software. Specifically, you want to find the book "Introduction to Quantum Mechanics" by David J. Griffiths. How will you find it?

You will use a **GET** request. First, you need to know how to write the query. Here are a couple of rules:

- space = "+"
- use "?q=" to start the query
- use "intitle:" to search for a word in the title
- use "inauthor:" to search for a word in the author name
- separate everything else with "+"

Okay, let's get started:

"?q=" to start the query <br>
"intitle:Introduction+intitle:to+intitle:Quantum+intitle:Mechanics" to search for our title <br>
"inauthor:David+inauthor:Griffiths" to search for our author

Putting them all together:

"?q=intitle:Introduction+intitle:to+intitle:Quantum+intitle:Mechanics+inauthor:David+inauthor:Griffiths"

Now that we have made our query, we are ready to make our **GET** request. Below is the code you will need to do this. In the top right corner, you can select the language which you prefer to use.

```{r get-example, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("code-chunk.html"), collapse="\n")
HTML(rawHTML)
```

If you are doing the Bash option, the results should print out automatically. For the other languages, we can print the response:

```{r get-example-2, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("code-chunk-get-2.html"), collapse="\n")
HTML(rawHTML)
```

In printing the results, the status code should be "200," and the printed text should be something like this:

```
{
 "kind": "books#volumes",
 "totalItems": 1,
 "items": [
  {
   "kind": "books#volume",
   "id": "0h-nDAAAQBAJ",
   "etag": "OTURZeInoRk",
   "selfLink": "https://www.googleapis.com/books/v1/volumes/0h-nDAAAQBAJ",
   "volumeInfo": {
    "title": "Introduction to Quantum Mechanics",
    "authors": [
     "David Griffiths"
    ],
    "publisher": "Cambridge University Press",
    "publishedDate": "2016-08-31",
    "description": "This best-selling textbook teaches students how to do quantum mechanics and provides an insightful discussion of what it actually means.",
    "industryIdentifiers": [
     {
      "type": "ISBN_13",
      "identifier": "9781107179868"
     },
     {
      "type": "ISBN_10",
      "identifier": "1107179866"
     }
    ],
    "readingModes": {
     "text": false,
     "image": true
    },
    "pageCount": 465,
    "printType": "BOOK",
    "categories": [
     "Science"
    ],
    "averageRating": 4.0,
    "ratingsCount": 4,
    "maturityRating": "NOT_MATURE",
    "allowAnonLogging": false,
    "contentVersion": "preview-1.0.0",
    "panelizationSummary": {
     "containsEpubBubbles": false,
     "containsImageBubbles": false
    },
    "imageLinks": {
     "smallThumbnail": "http://books.google.com/books/content?id=0h-nDAAAQBAJ&printsec=frontcover&img=1&zoom=5&edge=curl&source=gbs_api",
     "thumbnail": "http://books.google.com/books/content?id=0h-nDAAAQBAJ&printsec=frontcover&img=1&zoom=1&edge=curl&source=gbs_api"
    },
    "language": "en",
    "previewLink": "http://books.google.ca/books?id=0h-nDAAAQBAJ&printsec=frontcover&dq=intitle:Introduction+intitle:to+intitle:Quantum+intitle:Mechanics+inauthor:David+inauthor:Griffiths&hl=&cd=1&source=gbs_api",
    "infoLink": "http://books.google.ca/books?id=0h-nDAAAQBAJ&dq=intitle:Introduction+intitle:to+intitle:Quantum+intitle:Mechanics+inauthor:David+inauthor:Griffiths&hl=&source=gbs_api",
    "canonicalVolumeLink": "https://books.google.com/books/about/Introduction_to_Quantum_Mechanics.html?hl=&id=0h-nDAAAQBAJ"
   },
   "saleInfo": {
    "country": "CA",
    "saleability": "NOT_FOR_SALE",
    "isEbook": false
   },
   "accessInfo": {
    "country": "CA",
    "viewability": "PARTIAL",
    "embeddable": true,
    "publicDomain": false,
    "textToSpeechPermission": "ALLOWED",
    "epub": {
     "isAvailable": false
    },
    "pdf": {
     "isAvailable": false
    },
    "webReaderLink": "http://play.google.com/books/reader?id=0h-nDAAAQBAJ&hl=&printsec=frontcover&source=gbs_api",
    "accessViewStatus": "SAMPLE",
    "quoteSharingAllowed": false
   },
   "searchInfo": {
    "textSnippet": "This bestselling textbook teaches students how to do quantum mechanics and provides an insightful discussion of what it actually means."
   }
  }
 ]
}
```

Congrats! You have completed your first **GET** request!

### POST {-}

The **POST** call is a bit more complicated than the **GET** call. The primary difference is that for a **POST** call, you send data AND receive data. You probably have made a post on Facebook or Twitter or Instagram; those are examples of **POST** calls. Say you want to tweet "hello" on Twitter. You would send this data with a **POST** call to the Twitter API. The Twitter API would then add your new post, and will return a JSON object that looks something like this:

```
{
  "created_at": "Wed Oct 10 20:19:24 +0000 2018",
  "id": 1050118621198921700,
  "id_str": "1050118621198921728",
  "text": "To make room for more expression, we will now count all emojis as equal—including those with gender
and skin t… https://t.co/MkGjXf9aXm",
  "source": "<a href="http://twitter.com" rel="nofollow">Twitter Web Client</a>",
  "truncated": true,
  "in_reply_to_status_id": null,
  "in_reply_to_status_id_str": null,
  "in_reply_to_user_id": null,
  "in_reply_to_user_id_str": null,
  "in_reply_to_screen_name": null,
  "user": {
    "id": 6253282,
    "id_str": "6253282",
    "name": "Twitter API",
    "screen_name": "TwitterAPI",
    "location": "San Francisco, CA",
    "url": "https://developer.twitter.com",
    "description": "The Real Twitter API. Tweets about API changes, service issues and our Developer Platform. Don't get an answer? It's on my website.",
    "translator_type": "null",
    "derived": {
      "locations": [
        {
          "country": "United States",
          "country_code": "US",
          "locality": "San Francisco",
          "region": "California",
          "sub_region": "San Francisco County",
          "full_name": "San Francisco, California, United States",
          "geo": {
            "coordinates": [
              -122.41942,
              37.77493
            ],
            "type": "point"
          }
        }
      ]
    },
    "protected": false,
    "verified": true,
    "followers_count": 6172196,
    "friends_count": 12,
    "listed_count": 13003,
    "favourites_count": 31,
    "statuses_count": 3650,
    "created_at": "Wed May 23 06:01:13 +0000 2007",
    "utc_offset": null,
    "time_zone": null,
    "geo_enabled": false,
    "lang": "en",
    "contributors_enabled": false,
    "is_translator": null,
    "profile_background_color": "null",
    "profile_background_image_url": "null",
    "profile_background_image_url_https": "null",
    "profile_background_tile": null,
    "profile_link_color": "null",
    "profile_sidebar_border_color": "null",
    "profile_sidebar_fill_color": "null",
    "profile_text_color": "null",
    "profile_use_background_image": null,
    "profile_image_url": "null",
    "profile_image_url_https": "https://pbs.twimg.com/profile_images/942858479592554497/BbazLO9L_normal.jpg",
    "profile_banner_url": "https://pbs.twimg.com/profile_banners/6253282/1497491515",
    "default_profile": false,
    "default_profile_image": false,
    "following": null,
    "follow_request_sent": null,
    "notifications": null
  },
  "geo": null,
  "coordinates": null,
  "place": null,
  "contributors": null,
  "is_quote_status": false,
  "extended_tweet": {
    "full_text": "To make room for more expression, we will now count all emojis as equal—including those with gender




and skin tone modifiers. This is now reflected in Twitter-Text, our Open Source library. nnUsing Twitter-Text? See the forum post for detail: https://t.co/Nx1XZmRCXA",
    "display_text_range": [
      0,
      277
    ],
    "entities": {
      "hashtags": [],
      "urls": [
        {
          "url": "https://t.co/Nx1XZmRCXA",
          "expanded_url": "https://twittercommunity.com/t/new-update-to-the-twitter-text-library-emoji-character-count/114607",
          "display_url": "twittercommunity.com/t/new-update-t…",
          "unwound": {
            "url": "https://twittercommunity.com/t/new-update-to-the-twitter-text-library-emoji-character-count/114607",
            "status": 200,
            "title": "New update to the Twitter-Text library: Emoji character count",
            "description": "Over the years, we have made several updates to the way that people can communicate on Twitter. One of the more notable changes made last year was to increase the number of characters per Tweet from 140 to 280 characters. Today, we continue to expand people’s ability to express themselves by announcing a change to the way that we count emojis. Due to the differences in the way written text and emojis are encoded, many emojis (including emojis where you can apply gender and skin tone) have count..."
          },
          "indices": [
            254,
            277
          ]
        }
      ],
      "user_mentions": [],
      "symbols": []
    }
  },
  "quote_count": 0,
  "reply_count": 0,
  "retweet_count": 0,
  "favorite_count": 0,
  "entities": {
    "hashtags": [],
    "urls": [
      {
        "url": "https://t.co/MkGjXf9aXm",
        "expanded_url": "https://twitter.com/i/web/status/1050118621198921728",
        "display_url": "twitter.com/i/web/status/1…",
        "indices": [
          117,
          140
        ]
      }
    ],
    "user_mentions": [],
    "symbols": []
  },
  "favorited": false,
  "retweeted": false,
  "possibly_sensitive": false,
  "filter_level": "low",
  "lang": "en",
}
}
```

Now, for the purpose of this tutorial, we won't use the Twitter API or the Google Books API, because they require an authentication token to complete the **POST** requests. Basically, if you want to **POST** a tweet, Twitter needs to know that it is you asking to post on your account, and not someone else. However, getting an authentication token can take a bit of time, as you have to set up a developer account etc.

In data science, sometimes a **POST** request looks a bit different than the Twitter **POST** request. For example, let's say there is a machine learning algorithm which is able to tell you if a picture is a Cat or a Dog. This algorithm is on a Server somewhere else, and you want to use the API to check if your photo is a Cat or a Dog. This would also be an example of a **POST** request. You would send the data, which would be the picture, using a **POST** call. The API would communicate with the Server, and the algorithm would calculate if the picture was a Cat or a Dog. Then the algorithm would send back data, say a JSON file with the text "image: Cat". Now you know that your image is a Cat!

### PUT {-}

The **PUT** request is similar to **POST** in that you can send and receive data. However, the **PUT** request is used less often. The reason for this is what computer scientists and mathematicians call "idempotency." An action is **idempotent** if repeating the action results in the same end result. Still kind of complicated, right?

Let's do an example. The **POST** request is not idempotent. If you make a post on Twitter today, saying, "I love tacos!", and then later on you make another post that says "I love tacos!", both of those posts will stay on Twitter. The second post won't erase the first post [even though people already know you love tacos]. You can post "I love tacos!" as many times as you want, and Twitter will just keep adding it to your status.

On the other hand, the **PUT** request is idempotent. **PUT** is less often used because it means finding and overriding a specific resource. For example, say you have an online library, where a book is a resource. You would probably want to have a **GET** request, so people could get a book and read it. You might have a **POST** request, where people could add a new book to the library. A **PUT** request would be finding a book and replacing it with a newer version. So maybe you want to find the textbook "Introduction to Quantum Mechanics 2," and replace it with the latest version, "Introduction to Quantum Mechanics 3." You could use a **PUT** request for this. As you may have figured out, the **PUT** request can potentially delete items. Probably, you don't want just any person being able to replace books in the library. So, you might have a few "librarians" who are able to make **PUT** requests.

### DELETE {-}

**DELETE** is the last **REST** call, and it is fairly self-explanatory. In keeping with the online library, this would be deleting a book altogether. Similar to the **PUT** request, this generally is a call that only certain users have access to.

Another example of the **DELETE** request would be in Facebook. Maybe you made a post 5 years ago that was fairly embarassing. Maybe it was "I love tacos!" You could use a **DELETE** request to delete this post.

### Summary {-}

In **REST** jargon, we call data on the Server a **resource**. So in Google Books, the resource is a book. On Twitter, a resource might be a tweet. I wanted to point out here that a Server can have more than one resource, and can therefore have more than one **GET**/**POST**/**PUT**/**DELETE** calls. On the flip side, an API does not need to have all four calls. As you will see, often the only calls are **GET** and **POST**. In our API (PRISM), which you will read more about later, we only have **POST** calls at the moment.

I have taken a table from the Twitter API documentation to give you an example:

```{r twitter, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("twitter-post.html"), collapse="\n")
HTML(rawHTML)
```

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("link-buttons.html"), collapse="\n")
HTML(rawHTML)
```

