
# Using the PRISM API {#using-prism}

If you don't have much programming experience, or if you want to use Excel, I would recommend doing the [API in Excel](#api-in-excel). Otherwise, you can choose your favourite programming language to get started:

<!-- ```{r links, eval=TRUE, echo=FALSE} -->
<!-- library(htmltools) -->
<!-- rawHTML <- paste(readLines("link-buttons.html"), collapse="\n") -->
<!-- HTML(rawHTML) -->
<!-- ``` -->

## PRISM API in Excel {#api-in-excel}

## PRISM API in R/Python/Bash/Javascript {#api-in-R}

To get started, you will need to install a few packages:

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("install.html"), collapse="\n")
HTML(rawHTML)
```

The next step is to get an API key. For now, we will be giving out API keys on a case-by-case basis, so please email either:

Amin Adibi: amin.adibi@ubc.ca 

Mohsen Sadatsafavi: mohsen.sadatsafavi@ubc.ca

For the rest of this tutorial, I will be using one of the packages, **accept**, as the example. For a complete list of models, go to:

[PRISM Model Repository](http://resp.core.ubc.ca/ipress/prism)

### GET {-}

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("get-prism.html"), collapse="\n")
HTML(rawHTML)
```

### POST {-}

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("post-prism.html"), collapse="\n")
HTML(rawHTML)
```





